Entropy-SGD: Biasing Gradient Descent Into Wide Valleys

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
minor

In order to test the efficacy of $\entropysgd$ on recurrent architectures, we train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction. This dataset contains about one million words divided into a training set of about $930,000$ words, a validation set of $74,000$ words and $82,000$ words with a vocabulary of size $10,000$. As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}. Our network called $\ptblstm$ consists of two layers with $1500$ hidden units, each unrolled for $35$ time steps and a dropout of probability $0.65$; note that this is a large network with about $66$ million weights. It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch. \pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which closely matches the results of the original authors.

@@

Tuning the momentum in $\entropysgd$ was crucial to getting good results on RNNs. While the SGD baseline on $\ptblstm$ does not use momentum (and in fact, does not train well with momentum) we used a momentum of $0.5$ for $\entropysgd$. On the other hand, the baseline for $\charlstm$ was trained with Adam with $\b_1 = 0.9$ ($\b_1$ in Adam controls the moving average of the gradient) while we set $\b_1 = 0.5$ for $\entropyadam$. In contrast to this observation about RNNs, all our experiments on CNNs used a momentum of $0.9$. \pc{This is also corroborated by monitoring the angle between the local entropy gradient and the vanilla SGD gradient during training. This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former.}{}

@@

>>>>

In order to test the efficacy of $\entropysgd$ on recurrent architectures, we train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction. This dataset contains about one million words divided into a training set of about $930,000$ words, a validation set of $74,000$ words and $82,000$ words with a vocabulary of size $10,000$. As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}. Our network called $\ptblstm$ consists of two layers with $1500$ hidden units, each unrolled for $35$ time steps and a dropout of probability $0.65$; note that this is a large network with about $66$ million weights. It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch. We obtained a word perplexity of $81.43 \pm 0.2$ on the validation set and $78.6 \pm 0.26$ on the test set with this setup, which matches the results of the original authors very closely.

@@

Tuning the momentum in $\entropysgd$ was crucial to getting good results on RNNs. While the SGD baseline on $\ptblstm$ does not use momentum (and in fact, does not train well with momentum) we used a momentum of $0.5$ for $\entropysgd$. On the other hand, the baseline for $\charlstm$ was trained with Adam with $\b_1 = 0.9$ ($\b_1$ in Adam controls the moving average of the gradient) while we set $\b_1 = 0.5$ for $\entropyadam$. In contrast to this observation about RNNs, all our experiments on CNNs used a momentum of $0.9$. \pc{In order to investigate this difficulty, we monitored the angle between the local entropy gradient and the vanilla SGD gradient during training. This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy's gradient is thus highly uncorrelated with the SGD gradient in this case}{}

@@

- one paragraph on the efficacy of SGD and why it matters
